### 2024-04-15 paper/paper.qmd
update writings

---
title: "Mental Health Plays Significant Role in Physical Wellnesss in 1972 to 2022 US General Social Surveys"
subtitle: "Mental Illness Being Overlooked in the Medical Field with Racial and Gender Unequity"
author: 
  - Mingjia Chen
thanks: "Code and data are available at: https://github.com/MjChen120/Mental_to_Physical_Health.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(janitor,here,dplyr,tidyverse,knitr,kableExtra,ggplot2,arrow)

library(janitor)
library(tidyverse)
library(knitr)
library(dplyr)
library(kableExtra)
library(here)
library(ggplot2)
library(arrow)

#### Load data ####
GSS <- read_parquet(here::here("data/analysis_data/cleaned_GSS.parquet"))
demo_data <- read_parquet(here::here("data/analysis_data/demo_data.parquet"))
days_data <- read_parquet(here::here("data/analysis_data/days_data.parquet"))
analysis_data <- read_parquet(here::here("data/analysis_data/analysis_data.parquet"))
mentalVsHealth <- read_parquet(here::here("data/analysis_data/mentalVsHealth_data.parquet"))
```

# Introduction {#sec-intro}

Mental health is one essential aspect of the general health of an individual and yet it is often overlooked. Mental illnesses are not only “silent diseases” of one’s mind, but could as well impact patients’ physical well-being in general [@Ohrnberger_Fichera_Sutton]. This paper examines whether mental health-related and demographic factors impact the number of days an individual experiences physical unwellness in a month?

U.S. General Social Survey (GSS) data from 1972 to 2022 is used for this paper to examine whether the demographic and mental health-related factors as indicators of physical health conditions. The estimand is the difference in whether an individual could experience more occasions of feeling physically unwell if they have mental health unwellness versus not. This is considered in terms of those who were diagnosed as having depression by a doctor.

The analysis of GSS data, including gender, age, and mental and physical health variables, seeks to provide detailed insights into how mental health issues impact physical unwellness, contributing to discussions on intersectionality (complex sociological analytically framework for studying social and political identities resulted from unique combinations of discrimination and privilege) and overlooked mental health in the medical field.

<Results paragraph>

This paper begins with [Introduction] framing the impact of mental health on one's well-being. The [Data] section details variables, methodology, and analysis cleaning processes. In the [Model] section, a model is proposed for determining how an individual's depression status and counted days of mental unwellness in the past 30 days impact their counted days of physical unwellness in the past 30 days. [Results] section analyzes how demographic and other mental health-related factors would affect physical health with model calculations, leading the [Discussion] section with gender intersectionality and mental health issues often being overlooked in the medical field. Limitations and Future Research sections are included for study limitations and future research directions. The [Appendix] provides further survey information.

# Data {#sec-data}

The paper uses data collected from the US General Social Survey (GSS) from NORIC at the University of Chicago [@GSS]. From the dataset, this paper focuses specifically on variables related to demographic backgrounds, mental health, and physical health, from 1992 to 2022. This longitudinal approach allows us to analyze whether an individual's demographic factors and mental health well-being could significantly contribute to their physical well-being.

## Source Data

Physical and mental health variables such as "health", "depress" status, and reported days of physical ("physhlth") and health ("mntlhlth") unwellness in a month are selected from GSS [@GSS] for main analysis. Variable "year" is chosen for checking data consistency and non-response rate over the years. Demographic data, specifically age and gender, were used as well for further analyses. The data was downloaded and filtered for the selected variables from the selected data variables from GSS[^1]. The data cleaning was performed based on value definitions as defined in the GSS codebooks [@gssCodebook]. The variable names are renamed to be more informative (@tbl-use-var).

| Variable | New Name  | Description                   | Example |
|----------|-----------|-------------------------------|---------|
| ID_      | id        | Response ID                   | 1       |
| YEAR     | year      | Year of the Data Recorded     | 1977    |
| SEX      | sex       | Respondent's gender           | Female/2|
| AGE      | age       | Respondent's age              | 25      |
| HEALTH   | health    | Respondents' health condition | Good/2  |
| PHYSHLTH | phys_days | Days of Respondents' physical health being not good   | 15      |
| MNTLHLTH | ment_days | Days of Respondents' mental health being not good     | 20      |
| DEPRESS  | depress   | Whether respondents have been told having depression  | Yes/1   |

: Source data retrieved from GSS {#tbl-use-var}

## Data Cleaning

The data was cleaned by using the open source statistically programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], `arrow` [@arrow], and `knitr` [@knitr].

[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip

```{r}
#| label: tbl-sub-vars
#| tbl-cap: Overview of the Mental and Physical Unwellness
#| echo: false
#| warning: false
#| message: false

vars <- c("ID","Physical Unwellness","Mental Unwellness")

kable(head(days_data), 
      caption = "Counted days of Mental unwellness with General Health Status", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

## Survey Methodology

From 1992 until 2018, the General Social Survey (GSS) relied on face-to-face interviews to gather data. This method offered a robust data collection environment, allowing interviewers to delve deeper into responses and clarify any uncertainties. It provided an optimal means of gathering comprehensive data. However, in response to the COVID-19 Pandemic, between 2020 and 2021, the GSS transitioned primarily to web surveys for data collection.

The GSS targeted English or Spanish-speaking adults aged 18 and above residing in US households [@GSS]. This inclusive criteria aimed to attract a diverse pool of participants. Nevertheless, individuals falling outside this target demographic were deemed ineligible for the survey. Those who didn't speak English or Spanish or had mental and/or physical impediments preventing survey participation were considered outside the scope of the GSS [@GSS].

Changes in the GSS's respondent selection process, such as modifications to the Kish grid methodology, present additional hurdles, potentially skewing the demographic representation of respondents. The move towards web-based surveys may have inadvertently excluded older demographics, who are typically less inclined to engage with online platforms, thus impacting the survey's demographic balance. This demographic shift is significant, as older populations may offer distinct perspectives, particularly concerning mental health issues.

The survey queried participants with the questions, such as "Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?" and "Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good?" and expecting a response of number of days in a month. Significantly, between 1998 and 2022, the structure of the questions have consistently remained the same. This consistency in formatting ensures the data's stability, facilitating a high level of consistency for thorough analysis. The data consists of a combination of in-person, telephone, and web surveys.

## Demographic Variables

Demographic variables such as gender and age are used to compare and contrast differences between population groups. To further analyze age, the continuous variable is categorized into cohorts "0-19","20-39","40-59","60-79", and "80+". Below attached an overview of the sub-dataset used for demographic analysis.

```{r}
#| label: tbl-demo-overview
#| tbl-cap: Overview of the Demographic Data 
#| echo: false
#| warning: false
#| message: false

vars <- c("ID","Gender","Age","Physical Unwellness","Mental Unwellness","Age Cohort")

kable(head(demo_data), 
      caption = "Counted days of Physical and Mental unwellness of Population by Age and Gender Groups", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

## Mental and Physical Unwellness and Depression Diagnosis

One of the key variable used for further analysis is whether the respondent have been diagnosed as having depression by a doctor in the past. The survey used the question "Now I would like to ask you some questions about general health conditions. Has a doctor, nurse, or other health professional EVER told you that you had: D. Depression?" to collect a categorical response of "Yes" or "No". The variable is an external indication of respondents' mental health status, accompanying the main counted days variable that is used for the analysis.

```{r}
#| label: tbl-analysis-vars
#| tbl-cap: Overview of the Main Analysis Dataset
#| echo: false
#| warning: false
#| message: false

vars <- c("ID","Physical Unwellness","Mental Unwellness","Depression Diagnosis")

kable(head(analysis_data), 
      caption = "Counted days of Physical and Mental unwellness with Depression Diagnosis", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

## Mental Unwellness and General Health Status

A variable collecting general health status is furthermore included for additional analyses. The variable is used as a supplementary tool indicating respondents' health in general to assist the analysis result. The question consists of four categorical responses by asking the respondents: "Would you say your own health, in general, is excellent, good, fair, or  poor?"

```{r}
#| label: tbl-health-vars
#| tbl-cap: Overview of the Mental Unwellness vs. General Health Status Dataset 
#| echo: false
#| warning: false
#| message: false

vars <- c("ID","Mental Unwellness","General Health")

kable(head(mentalVsHealth), 
      caption = "Counted days of Mental unwellness with General Health Status", 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

## Non-response Rate

To mitigate non-responses in the survey, the GSS implemented a technique known as subsampling for non-respondents. The core concept of subsampling entails selecting a subset of non-respondents and adjusting their weights to ensure an impartial design [@gssCodebook2014]. This chosen subsample is then expanded to effectively mirror all non-respondents up to a predetermined cutoff date. By focusing resources on a more manageable subset of difficult cases for further follow-up, subsampling offers the potential to reduce both response error and non-response bias [@gssCodebook2014].

The GSS's approach to managing non-responses underwent changes, notably in its recent iterations – specifically during the COVID-19 Pandemic. As web surveys were used in the 2021 survey, modifications were made to balance non-responses. Historically, the survey employed subsampling techniques to mitigate the impact of non-responses, ensuring a representative dataset.

However, with the introduction of web-based surveys, new categories like "Skipped on Web" were introduced to categorize non-responses. This change raises questions about the comparability of data across years and the potential for increased non-response rates. Non-responses (@fig-nonresponse-rate) can significantly affect the analysis of physical and mental health status trends, particularly in capturing the full extent of respondents' health status. Furthermore, as mental health is a later introduced science compared to physical health and thus is not included in most of the previous surveys. Although survey data lack of mental health related questions is filtered out for analyses, it is important to take that fact into consideration.

```{r}
#| label: fig-nonresponse-rate
#| fig-cap: Nonresponse Rates
#| echo: false
#| warning: false
#| message: false

# record number of non-responses in variables ment_days, phys_days, and depress
no_response_data <- rep(0,3)

#length(which(<condition>)) code referenced from https://www.geeksforgeeks.org/count-repeated-values-in-r/
no_response_data <- c(length(which(GSS$phys_days<1)),length(which(GSS$ment_days<1)),length(which(GSS$depress<1)))

# calculate the percentages
total_responses <- nrow(GSS)
non_response_rate <- no_response_data/total_responses 
non_response_rate <- round(non_response_rate,3)
questions <- c("Physical Unwellness","Mental Unwellness","Depression Status")
non_response_data <- cbind(questions,non_response_rate)

ggplot(non_response_data, aes(x = questions,y=non_response_rate)) +
  geom_bar(stat = "identity",fill = "skyblue") +
  theme_minimal() +
  labs(title = "Non-Response Rates of each Question", x = "Questions", y = "Percentage of Non-Responses")

```

As shown in @fig-nonresponse-rate, some of the questions in the dataset are not included resulting no responses collected in certain years due to the change in survey methodology. This fact requires extra attention when cleaning and handling since the number of available responses will decrease when additional variables are added. Hence, an additional sub set of the data that only includes counted days of mental and physical un-wellness in the past 30 days (@tbl-sub-vars) is also included for only speculating these two variables while keeping as many responses as possible.

\break

# Model
Through the data analysis of the paper, a potential correlation between counted days of feeling mental and physical unwell is detected. To further analyze and estimate counted days of feeling physical unwell based on an individuals counted days of feeling mentally unwell and whether they were diagnosed as depressive in the past, a linear regression analysis is conducted.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of days that an individual feel physically unwell in the past 30 days. Then $\beta_i$ is the number of days that this individual feel mentally unwell and $\gamma_i$ indicates whether the individual is diagnosed as depressive by a doctor. $y_i$ and $\beta_i$ are measured in days.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

will be written

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| warning: false
#| message: false

#categorize the range of days for visualizing the histogram of demographic variables

demo_data <- within(demo_data, {   
  ment_range <- NA # need to initialize variable
  ment_range[ment_days < 10] <- "0-10"
  ment_range[ment_days >= 10 & ment_days < 20] <- "10-20"
  ment_range[ment_days >= 20 & ment_days < 30] <- "20-30"
} )

demo_data <- within(demo_data, {   
  phys_range <- NA # need to initialize variable
  phys_range[phys_days < 10] <- "0-10"
  phys_range[phys_days >= 10 & phys_days < 20] <- "10-20"
  phys_range[phys_days >= 20 & phys_days < 30] <- "20-30"
} )
```

```{r}
#| label: fig-demo-mental
#| fig-cap: Female Participants from 20 to 59 Years Old are More likely to Experience Mental Unwellness 
#| fig-subcap: ["Gender","Age cohorts"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

demo_data |>
  ggplot(aes(x = ment_days, fill = sex)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Mentally in Last 30 days",
    y = "Occurrences",
    fill = "Sex"
  )

demo_data |>
  ggplot(aes(x = ment_days, fill = age_cohort)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Mentally in Last 30 days",
    y = "Occurrences",
    fill = "Age Cohorts"
  )

```
\break

```{r}
#| label: fig-demo-physical
#| fig-cap: Female Participants from 20 to 59 Years Old are More likely to Experience Physical Unwellness 
#| fig-subcap: ["Gender","Age cohorts"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

demo_data |>
  ggplot(aes(x = phys_days, fill = sex)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Physically in Last 30 days",
    y = "Occurrences",
    fill = "Sex"
  )

demo_data |>
  ggplot(aes(x = phys_days, fill = age_cohort)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Physically in Last 30 days",
    y = "Occurrences",
    fill = "Age Cohorts"
  )

```
\break

Include some research about female having trouble due to medical racism etc. and possible reason why these age cohorts are more likely to experience both physical and mental unwellness.

## Mental Health and Health in General
```{r}
#| label: fig-mental-vs-general
#| fig-cap: Counts of Days Feeling Unwell in Last 30 Days
#| echo: false
#| warning: false
#| message: false

#Graph
mentalVsHealth|>
  ggplot(aes(x = ment_days, y = health)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Mental Unwellness", y = "General Health") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```
\break

## Depression Playing an Role
```{r}
#| label: fig-mental-vs-physical
#| fig-cap: Counts of Days Feeling Unwell in Last 30 Days
#| echo: false
#| warning: false
#| message: false

#Graph
analysis_data|>
  ggplot(aes(x = ment_days, y = phys_days, color = depress)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Mental Unwellness", y = "Physical Unwellness", color = "Depress") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```
\break

## Model Results

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of counted days of physical unwellness based on counted days of mental unwellness and depression diagnosis"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix

## Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\break

\newpage

# References