### 2024-04-23 paper/paper.qmd
final check + fix typos

---
title: "Mental Health Plays Significant Role in Physical Wellness From 1972 to 2022 US General Social Surveys"
subtitle: "Mental Illness Being Overlooked in the Medical Field with Gender Disparities"
author: 
  - Mingjia Chen
thanks: "Code and data are available at: https://github.com/MjChen120/Mental_to_Physical_Health.git."
date: today
date-format: long
abstract: "Mental health is a relatively newly introduced topic in the field, but would mental health issues influence physical health as well? Focusing on numerous variables correlated to mental and physical health, this study uses data from the U.S. General Social Survey (GSS) from 1972 to 2022 to investigate the impact of mental health unwellness on physical health unwellness, focusing specifically on “reported days of feeling unwell in a month.” Our analysis reveals that mental unwellness leads to an increase in physical unwellness, and identifies disparities between demographic groups, with women disproportionately affected. These findings underscore the necessity for the importance of promoting mental health and supporting underrepresented groups in the medical field."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(janitor, here, dplyr, tidyverse, knitr, kableExtra, ggplot2, arrow, ggpmisc)

library(janitor)
library(tidyverse)
library(knitr)
library(dplyr)
library(kableExtra)
library(here)
library(ggplot2)
library(arrow)
library(ggpmisc)

#### Load data ####
GSS <- read_parquet(here::here("data/analysis_data/cleaned_GSS.parquet"))
demo_data <- read_parquet(here::here("data/analysis_data/demo_data.parquet"))
days_data <- read_parquet(here::here("data/analysis_data/days_data.parquet"))
analysis_data <- read_parquet(here::here("data/analysis_data/analysis_data.parquet"))
mentalVsHealth <- read_parquet(here::here("data/analysis_data/mentalVsHealth_data.parquet"))
```

# Introduction {#sec-intro}

Mental health is one essential aspect of the general health of an individual and yet it is often overlooked. Mental illnesses are not only “silent diseases” of one’s mind, but could as well impact patients’ physical well-being in general [@Ohrnberger_Fichera_Sutton]. Further research analyses should be conducted to benefit people's health with statistical information, thus it is important to speculate the relationship in depth. Even though previous research evidence supports the relationship between mental and physical health, larger representative samples and additional health-related variables were not used. To contribute to the topic, this paper examines whether mental health-related and demographic factors impact the number of days an individual experiences physical unwellness in a month by using the U.S. General Social Survey (GSS) dataset.

The estimand of the paper is the difference in whether an individual could experience more occasions of feeling physically unwell if they have mental health unwellness versus not. This is considered in terms of those who were diagnosed as having depression by a doctor. U.S. General Social Survey (GSS) data from 1972 to 2022 is therefore used for this paper to examine whether the demographic and mental health-related factors as indicators of physical health conditions.

The analysis of GSS data, including gender, age, and mental and physical health variables, seeks to provide detailed insights into how mental health issues impact physical unwellness, contributing to discussions on intersectionality (complex sociological analytical framework for studying social and political identities resulted from unique combinations of discrimination and privilege) and overlooked mental health in the medical field. The paper is essential to contribute to the narrative of the relationship between mental and physical health with the U.S. population dataset.

Findings in the paper highlight the significance of gender, sex, and counted days of feeling mentally unwell in terms of counted days of feeling physically unwell in a month. The effect of being diagnosed as having depression on physical health unwellness is not significant ($\beta_2$ = −1.43, p = .22), thus whether depression diagnosis affects physical unwellness is not supported by evidence. On the other hand, counted days of feeling mentally unwell have a significant positive influence on counted days of feeling physically unwell ($\beta_1$ = 0.27, p = .01). This means that if an individual has 1 more day of feeling unwell mentally, they, in general, will report 0.27 more days when reporting the number of days they feel physically unwell.

This paper begins with [Introduction] framing the impact of mental health on one's well-being. The [Data] section details variables, methodology, and analysis cleaning processes. In the [Model] section, a model is proposed for determining how an individual's depression status and counted days of mental unwellness in the past 30 days impact their counted days of physical unwellness in the past 30 days. [Results] section analyzes how demographic and other mental health-related factors would affect physical health with model calculations, leading the [Discussion] section with gender intersectionality and mental health issues often being overlooked in the medical field. Limitations and Future Research sections are included for study limitations and future research directions. The [Appendix] provides further survey information.

# Data {#sec-data}

The paper uses data collected from the US General Social Survey (GSS) from the National Opinion Research Center (NORC) at the University of Chicago [@GSS]. From the dataset, this paper focuses specifically on variables related to demographic backgrounds, mental health, and physical health, from 1972 to 2022. This longitudinal approach allows us to analyze whether an individual's demographic factors and mental health well-being could significantly contribute to their physical well-being.

## Source Data

The source data was obtained from the public open website GSS Data Explorer[^1]. The data was downloaded and filtered for the selected variables and further cleaning and relabeling was performed based on value definitions as defined in the GSS codebooks [@gssCodebook]. The variable names are renamed to be more informative (@tbl-use-var). Physical and mental health variables such as "health", "depress" status, and reported days of feeling physically ("physhlth") and health ("mntlhlth") unwell in a month are selected from GSS [@GSS] for the main analysis. Variable "year" is chosen for checking data consistency and non-response rate over the years. Demographic data, specifically age and gender, were used as well for further analyses.

| Variable | New Name  | Description                   | Example |
|----------|-----------|-------------------------------|---------|
| ID_      | id        | Response ID                   | 1       |
| YEAR     | year      | Year of the Data Recorded     | 1977    |
| SEX      | sex       | Respondent's gender           | Female/2|
| AGE      | age       | Respondent's age              | 25      |
| HEALTH   | health    | Respondents' health condition | Good/2  |
| PHYSHLTH | phys_days | Days of Respondents' physical health being not good   | 15      |
| MNTLHLTH | ment_days | Days of Respondents' mental health being not good     | 20      |
| DEPRESS  | depress   | Whether respondents have been told having depression  | Yes/1   |

: Source data retrieved from GSS Data Explorer {#tbl-use-var}

[^1]: https://gssdataexplorer.norc.org/home

## Data Cleaning

The explained data was cleaned by using the open source statistical programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `ggpmisc` [@ggpmisc], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], `arrow` [@arrow], and `knitr` [@knitr]. Below in @tbl-sub-vars contains a snapshot of the main dataset being used for analysis; two additional histograms (@fig-countedDays) are included as well, illustrating the number of responses for each question. In general, most respondents have only a few or zero days of feeling both physically and mentally unwell in the past 30 days, indicating the population is generally well in mental and physical health. Very few responses were collected that fall in the ranges between 15 days to 20 days, and above 20 days and below 30 days. Remarkably, on the other hand, there are a certain amount of responses reporting 30 days, indicating percentages of participants experiencing high levels of health discomfort. The trends are similar between the two responses.

```{r}
#| label: tbl-sub-vars
#| tbl-cap: Counted days of Mental unwellness with General Health Status
#| echo: false
#| warning: false
#| message: false

vars <- c("ID", "Physical Unwellness", "Mental Unwellness")

kable(head(days_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

```{r}
#| label: fig-countedDays
#| fig-cap: Histograms for Number of Responses for Each Question
#| fig-subcap: ["Mental Unwellness", "Physical Unwellness"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

demo_data |>
  ggplot(aes(x = ment_days)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Mentally in Last 30 days",
    y = "Occurrences",
    fill = "Sex"
  )

demo_data |>
  ggplot(aes(x = phys_days)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Physically in Last 30 days",
    y = "Occurrences",
    fill = "Age Cohorts"
  )

```

## Survey Methodology

From 1972 until 2018, the General Social Survey (GSS) relied on face-to-face interviews to gather data. This method offered a robust data collection environment, allowing interviewers to delve deeper into responses and clarify any uncertainties. It provided an optimal means of gathering comprehensive data. However, in response to the COVID-19 Pandemic, between 2020 and 2021, the GSS transitioned primarily to web surveys for data collection.

The GSS targeted English or Spanish-speaking adults aged 18 and above residing in US households [@GSS]. This inclusive criteria aimed to attract a diverse pool of participants. Nevertheless, individuals falling outside this target demographic were deemed ineligible for the survey. Those who didn't speak English or Spanish or had mental and/or physical impediments preventing survey participation were considered outside the scope of the GSS [@GSS].

Changes in the GSS's respondent selection process, such as modifications to the Kish grid methodology, present additional hurdles, potentially skewing the demographic representation of respondents. The move towards web-based surveys may have inadvertently excluded older demographics, who are typically less inclined to engage with online platforms, thus impacting the survey's demographic balance. This demographic shift is significant, as older populations may offer distinct perspectives, particularly concerning mental health issues.

The survey queried participants with questions, such as "Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?" and "Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good?" and expecting a response of the number of days in a month. Further details of the questions used are listed in [Appendix -@sec-ques-details]. Significantly, between 1998 and 2022, the structure of the questions has consistently remained the same. This consistency in formatting ensures the data's stability, facilitating a high level of consistency for thorough analysis. The data consists of a combination of in-person, telephone, and web surveys.

## Demographic Variables

Demographic variables such as gender and age are used to compare and contrast differences between population groups. To further analyze age, as showcased in @tbl-demo-overview, the continuous variable is modified to be categorized into cohorts "0-19", "20-39", "40-59", "60-79", and "80+". Below is an overview of the sub-dataset used for demographic analysis. In total, as shown in @tbl-demoportion, 9049 responses (with approximately 51% from female respondents and 44% from age cohort 40 to 59) containing mentioned variables were collected. Age cohort from 20 to 39 takes 43% of the responses; in other words, most respondents are from age 20 to 59 (87%).

```{r}
#| label: tbl-demo-overview
#| tbl-cap: Counted days of Physical and Mental unwellness of Population by Age and Gender Groups
#| echo: false
#| warning: false
#| message: false

vars <- c("ID", "Gender", "Age", "Physical Unwellness", "Mental Unwellness", "Age Cohort")

kable(head(demo_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)
```

\break

```{r}
#| label: tbl-demoportion
#| tbl-cap: Number of Responses from Each Demographic Group
#| tbl-subcap: 
#|   - "Sex"
#|   - "Age Cohorts"
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

genderdata <- filter(demo_data, sex == "Female")

sex <- c("Male", "Female", "Total")
numbers <- c(nrow(genderdata), nrow(demo_data)-nrow(genderdata), nrow(demo_data))
female_per <- round(nrow(genderdata)/nrow(demo_data),2) * 100

gendertable <- cbind(sex,numbers)
kable(gendertable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Sex", "Responses"))

agetable <- count(demo_data, age_cohort)
total <- c("Total",nrow(demo_data))
agetable <- rbind(agetable,total)
kable(agetable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Age Cohort", "Responses"))

age_per <- round(nrow(filter(demo_data,age_cohort == "40-59"))/nrow(demo_data), 2) * 100
```

\break

## Mental and Physical Unwellness and Depression Diagnosis

One of the key variables used for further analysis is whether the respondent has been diagnosed as having depression in the past. The survey used the question "Now I would like to ask you some questions about general health conditions. Has a doctor, nurse, or other health professional EVER told you that you had: D. Depression?" to collect a categorical response of whether "Yes" or "No". The variable is an external indication of respondents' mental health status, accompanying the main counted days variable that is used for the analysis (@tbl-analysis-vars). From @tbl-analysis-depress, about 21% of the participants have been diagnosed as having depression in the past. According to the Morbidity and Mortality Weekly Report from the Centers for Disease Control and Prevention (CDC) in the U.S.[@CDControlPrevention_2023], 18.4% of U.S. adults reported having ever been diagnosed with depression in 2020. The percentage in dataset is slightly higher than weekly reports, which could be a result of error, sampling differences, and other factors.

```{r}
#| label: tbl-analysis-vars
#| tbl-cap: Counted days of Physical and Mental unwellness with Depression Diagnosis
#| echo: false
#| warning: false
#| message: false

vars <- c("ID", "Physical Unwellness", "Mental Unwellness", "Depression Diagnosis")

kable(head(analysis_data), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)

per <- round(count(filter(analysis_data, depress=="Yes")) / count(analysis_data), 2) * 100

```
\break
```{r}
#| label: tbl-analysis-depress
#| tbl-cap: Distribution of Depression Diagnosis Responses
#| echo: false
#| warning: false
#| message: false

depresstable <- count(analysis_data, depress)
total <- c("Total",nrow(analysis_data))
depresstable <- rbind(depresstable, total)
kable(depresstable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Responses", "Occurrences"))
```

## Mental Unwellness and General Health Status

A variable collecting general health status is furthermore included for additional analyses (@tbl-health-vars). The variable is used as a supplementary tool indicating respondents' health in general to assist the analysis result. The question consists of four categorical responses asking the respondents: "Would you say your own health, in general, is excellent, good, fair, or poor?" The respondents are expected to choose one of the options to answer. In @tbl-health-rep, the number of occurrences for each response is counted. The majority of the respondents reported to have good (51%) or excellent (29%) health status.

```{r}
#| label: tbl-health-vars
#| tbl-cap: Overview of the Mental Unwellness vs. General Health Status Dataset 
#| echo: false
#| warning: false
#| message: false

per_good <- round(nrow(filter(mentalVsHealth, health == 2))/nrow(mentalVsHealth),2) * 100
per_exc <- round(nrow(filter(mentalVsHealth, health == 1))/nrow(mentalVsHealth),2) * 100

vars <- c("ID", "Mental Unwellness", "General Health")
kable(head(mentalVsHealth), 
      align = "c",
      col.names = vars,
      row.names = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)

```

```{r}
#| label: tbl-health-rep
#| tbl-cap: Distribution of General Health Responses
#| echo: false
#| warning: false
#| message: false
healthtable <- count(mentalVsHealth, health)
total <- c("Total", nrow(mentalVsHealth))
healthtable <- rbind(healthtable, total)
status <- data.frame(c("Excellent", "Good", "Fair", "Poor","Total"))
colnames(status) <- c("Health Status")
healthtable <- cbind(status, healthtable)
kable(healthtable, 
      align = "c",
      row.names = FALSE,
      col.names = c("Health Status", "Responses", "Occurrences"))
```

## Non-response Rate and Limitations

To mitigate non-responses in the survey, the GSS implemented a technique known as subsampling for non-respondents. The core concept of subsampling entails selecting a subset of non-respondents and adjusting their weights to ensure an impartial design [@gssCodebook2014]. This chosen subsample is then expanded to effectively mirror all non-respondents up to a predetermined cutoff date. By focusing resources on a more manageable subset of difficult cases for further follow-up, subsampling offers the potential to reduce both response error and non-response bias [@gssCodebook2014].

The GSS's approach to managing non-responses underwent changes, notably in its recent iterations – specifically during the COVID-19 Pandemic. As web surveys were used in the 2021 survey, modifications were made to balance non-responses. Historically, the survey employed subsampling techniques to mitigate the impact of non-responses, ensuring a representative dataset.

However, with the introduction of web-based surveys, new categories like "Skipped on Web" were introduced to categorize non-responses. This change raises questions about the comparability of data across years and the potential for increased non-response rates. Non-responses (@fig-nonresponse-rate) can significantly affect the analysis of physical and mental health status trends, particularly in capturing the full extent of respondents' health status. Furthermore, mental health is a later introduced science compared to physical health and thus is not included in most of the previous surveys. Although survey data lack mental health-related questions filtered out for analyses, it is important to consider that fact.

As shown in @fig-nonresponse-rate, some of the questions in the dataset are not included resulting in no responses collected in certain years due to the change in survey methodology. This fact requires extra attention when cleaning and handling since the number of available responses will decrease when additional variables are added. Hence, an additional subset of the data that only includes counted days of mental and physical un-wellness in the past 30 days (@tbl-sub-vars) is also included for only speculating these two variables while keeping as many responses as possible.

```{r}
#| label: fig-nonresponse-rate
#| fig-cap: Nonresponse Rates - Responses for counted days of feeling physically unwell have the highest nonresponse rate
#| echo: false
#| warning: false
#| message: false

# record number of non-responses in variables ment_days, phys_days, and depress
no_response_data <- rep(0,3)

#length(which(<condition>)) code referenced from https://www.geeksforgeeks.org/count-repeated-values-in-r/
no_response_data <- c(length(which(GSS$phys_days<1)), length(which(GSS$ment_days<1)), length(which(GSS$depress<1)))

# calculate the percentages
total_responses <- nrow(GSS)
non_response_rate <- no_response_data/total_responses 
non_response_rate <- round(non_response_rate, 3)
questions <- c("Physical Unwellness", "Mental Unwellness", "Depression Status")
non_response_data <- cbind(questions, non_response_rate)

ggplot(non_response_data, aes(x = questions, y=non_response_rate)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(x = "Questions", y = "Proportion of Non-Responses")

```
\break

# Model

Through the data analysis of the paper, a potential correlation between counted days of feeling mentally and physically unwell is detected. To further analyze and estimate counted days of feeling physically unwell based on an individual's counted days of feeling mentally unwell and whether they were diagnosed as depressive in the past, a linear regression analysis is conducted.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of days that an individual felt physically unwell in the past 30 days. Then $\beta_1$ is the coefficient for the number of days that this individual felt mentally unwell and $\beta_2$ indicates the coefficient of whether the individual is diagnosed as depressive by a doctor. $y_i$ and days are measured in days.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 days + \beta_2 diagnosis\\
\beta_i &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`. Model justification and result will be explained later in this paper.

## Model justification

We anticipate that the counted days of feeling physically unwell would be closer to a normal distribution, as different types of health-related factors would contribute to number of days people feeling unwell in a month; however, the relationship between days feeling unwell physically (dependent variable), days feeling unwell mentally (independent variable), and whether being diagnosed as having depression before (independent variable) is hypothesized to be linear. This is because we think feeling mentally unwell and having depression would contribute to a higher possibility of feeling unwell physically, corresponding to our estimand. Since these three questions were not included in all past surveys, only data that contains all three questions will be used for testifying the model.

# Results {#sec-results}

## Differences Detected between Demographic Populations

To speculate possible demographic factors influencing days of feeling mentally and physically unwell in a month, gender and sex are used. In general, most of the participants reported having less than 10 days of feeling unwell both mentally and physically. Throughout the responses except for 0 days, the occurrence of responses from female participants is higher than from male participants; this means that a larger percentage of female participants have experienced both mentally and physically unwell in the past 30 days. Age cohorts from 20 to 59 contribute to the majority of the occurrences for each response, which could be a result of the fact that the age cohort takes up to 87% of the dataset. In conclusion, females from 20 to 39 years old are more likely to report a higher number of days feeling mentally unwell (@fig-demo-mental). In terms of physical health, females from 40 to 59 years old report a higher number of days of feeling unwell (@fig-demo-physical).

As we can see, females in general report a higher number of counted days of feeling both physically and mentally unwellness in a month. The difference could be disparity occurring between gender groups, such as females usually being less represented in medical research. On the other hand, females are reported to be more sensitive to their mental health, which could be the reason why they tend to have a higher number of reports of being mentally unwell.

```{r}
#| echo: false
#| warning: false
#| message: false

#categorize the range of days for visualizing the histogram of demographic variables

demo_data <- within(demo_data, {   
  ment_range <- NA # need to initialize variable
  ment_range[ment_days < 10] <- "0-10"
  ment_range[ment_days >= 10 & ment_days < 20] <- "10-20"
  ment_range[ment_days >= 20 & ment_days < 30] <- "20-30"
} )

demo_data <- within(demo_data, {   
  phys_range <- NA # need to initialize variable
  phys_range[phys_days < 10] <- "0-10"
  phys_range[phys_days >= 10 & phys_days < 20] <- "10-20"
  phys_range[phys_days >= 20 & phys_days < 30] <- "20-30"
} )
```

```{r}
#| label: fig-demo-mental
#| fig-cap: Female Participants from 20 to 39 Years Old are More Likely to Experience Mental Unwellness 
#| fig-subcap: ["Gender", "Age cohorts"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

demo_data |>
  ggplot(aes(x = ment_days, fill = sex)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Mentally in Last 30 days",
    y = "Occurrences",
    fill = "Sex"
  )

demo_data |>
  ggplot(aes(x = ment_days, fill = age_cohort)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Mentally in Last 30 days",
    y = "Occurrences",
    fill = "Age Cohorts"
  )

```

```{r}
#| label: fig-demo-physical
#| fig-cap: Female Participants from 40 to 59 Years Old are More Likely to Experience Physical Unwellness 
#| fig-subcap: ["Gender", "Age cohorts"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

demo_data |>
  ggplot(aes(x = phys_days, fill = sex)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Physically in Last 30 days",
    y = "Occurrences",
    fill = "Sex"
  )

demo_data |>
  ggplot(aes(x = phys_days, fill = age_cohort)) +
  geom_histogram(position = "dodge") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    x = "Days of Feeling Unwell Physically in Last 30 days",
    y = "Occurrences",
    fill = "Age Cohorts"
  )

```

## Nonresponse Rate

As we illustrated in @fig-nonresponse-rate, non-response rates may also help indicate significant differences in participants' attitudes toward each question included in the survey. Physical health-related questions have the highest non-response rate, indicating that most of the participants lack the motivation to time on speculate their physical health. On the other hand, the non-response rate for mental health and depression diagnosis is lower than for physical health questions. This difference between non-response rates could show that participants are more likely to pay attention when they are feeling mentally unwell and are more willing to answer questions corresponding to mental health.

## Mental Health and Health in General

In addition to model and demographic analyses, a variable indicating respondents' general health was also included for supplementary support of the paper. The general health variable is categorical, consisting of four responses: "1" for Excellent, "2" for Good, "3" for Fair, and "4" for Poor. As shown in @fig-mental-vs-general, we can see that the number of days feeling mentally unwell has a positive correlation with the general health status, meaning that if an individual has more days feeling mentally unwell, they are more likely to have poor general health as well. This result connects back to our estimand, supporting mental health's influence on one's physical health status. The relationship, however, is not strong ($\beta_1 =0.03$, $R^2 = 0.07$); this could be due to the fact that the health variable was not framed to be correlated to mental health. The variable may fail to associate with other health variables.

```{r}
#| label: fig-mental-vs-general
#| fig-cap: Counts of Days Feeling Mentally Unwell in Last 30 Days Has Positive Correlation with Poor General Health.
#| echo: false
#| warning: false
#| message: false

#Graph
mentalVsHealth|>
  ggplot(aes(x = ment_days, y = health)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_minimal() +
  labs(x = "Mental Unwellness", y = "General Health") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```
\break

## Depression Playing a Role

With linear regression model analysis, the result in @fig-mental-vs-physical shows that counted days of feeling mentally unwell are positively correlated to counted days of feeling physically unwell. The diagnosis of depression has a further positive effect on counted days of feeling physically unwell, making the correlation stronger. This result means that individuals with 1) higher counted days of feeling mentally unwell and 2) being diagnosed as having depression tend to report a higher number of days feeling physically unwell.

Based on @tbl-modelresults, however, although having influences on linear equation in @fig-mental-vs-physical, depression diagnosis's influences on counted days of feeling physically unwell is not significant ($\beta_2$ = −1.43, p = .22). Counted days of feeling mentally unwell, on the other hand, do have a significant positive correlation with counted days of feeling physically unwell ($\beta_1$ = 0.27, p = .01). From @fig-mental-vs-physical, it is evident to state that depression diagnosis may be a moderator to strengthen the relationship between counted ays of mental and physical unwellness in the past 30 days. The lack of significance could be due to possible errors, validity of survey questions, biases, or other possible factors. Hence, even though one of the correlations is not significant, it is still important for us to consider the variable for possible future research.

```{r}
#| label: fig-mental-vs-physical
#| fig-cap: Counts of Days Feeling Mentally Unwell in Last 30 Days Has Positive Correlation with Counted Days of Feeling Physically Unwell, with Depression Diagnosis Affecting the Strength of the Correlation.
#| echo: false
#| warning: false
#| message: false

#Graph
analysis_data|>
  ggplot(aes(x = ment_days, y = phys_days, color = depress)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_poly_line() +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  theme_minimal() +
  labs(x = "Mental Unwellness", y = "Physical Unwellness", color = "Depress") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of counted days of physical unwellness based on counted days of mental unwellness and depression diagnosis"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

\break

# Discussion {#sec-discussion}

For the main analysis, we find a positive correlation between mental and physical health, where depression diagnosis strengthens the correlation; on the other hand, we detect a disparity between gender groups, hinting at an association with the fact that females are underrepresented in the medical field.

## Mental Health Issues Impact Physical Health {#sec-first-point}

Corresponding to the results in the journal by @Ohrnberger_Fichera_Sutton, although not strong, the result of the paper supports the positive correlation between mental and physical health status ($\beta_1$ = 0.27, p = .01). Many possible reasons can contribute to the result we have for the paper. The variables that are used for the main analysis were only available in the years 2002, 2006, 2010, 2014, 2018, and 2022 since mental health is a newly introduced concept in health science compared to physical health, hence the sample size (n = 9049) of our data may not accurately represent U.S. population in past decades. Moreover, mental health education may not be common, and even mistreated as a taboo, to the society at the time, causing individuals to dismiss their mental discomfort or issues. Participants might refuse to answer relevant questions because they are afraid of discrimination as those personal health variables could make them vulnerable to social rejection. The correlation or impact thus could be strengthened by mental health promotions and education in the future: having a society that promotes and educates mental health, participants have better knowledge of their own health status and more responses recollecting the accurate status will be available for analysis.

In addition, counted days of feeling physically and mentally unwell might not fully capture an individual's health status for analysis, more relevant factors should be included for future analysis. Due to that reason, a variable capturing respondents' general health status was included for supplementary evidence. A positive correlation between counted days of mental unwellness and general health (@fig-mental-vs-general) was detected to further support our estimand. To elaborate, since the categorical variable is coded reversely (i.e. Excellent - 1 and Poor - 4), we can conclude that as the number of counted mental unwellness increases, the health status of that individual correspondingly worsens. Keep in mind that the strength of the correlation is not strong, but the result provides insight into how mental health issues could influence health in general.

## Influence by Depression Diagnosis

Research [@Trivedi_the_Link_2004] indicates that physical symptoms frequently accompany depression, with vague aches and pains often being the primary indicators. These symptoms commonly manifest as chronic joint pain, limb discomfort, backaches, gastrointestinal issues, fatigue, sleep disturbances, alterations in psychomotor activity, and changes in appetite. A significant portion of patients initially seek medical attention solely for physical complaints, only to later discover a diagnosis of depression. Research provides an insight that mental health issues are often associated with physical symptoms. Thus, to support our estimand, an additional variable was added to indicate the respondents' depression diagnosis. From the graph (@fig-mental-vs-physical) and model (@tbl-modelresults) results, depression diagnosis contributes to increasing the influence of mental unwellness to physical unwellness. Depression diagnosis, therefore, could be a moderator, which strengthens the relation between mental health (independent variable) and physical health (dependent variable). Other than depression diagnosis, no additional mental health issues are speculated for the purpose of this particular study. The result with the GSS dataset, although not significant, supports the fact that depression plays a role in impacting physical health.

## Gender Disparity in Health
By comparing counted days of feeling mentally and physically unwell between different gender groups, we found that female respondents in general reported higher numbers of occasions feeling unwell (@fig-demo-mental and @fig-demo-physical). The trend could be a result of gender disparity in the medical field, where women benefit less from medical research and advances because they are underrepresented as potential patients. The gender disparity in medical research, compounded by pervasive misogyny, translates into tangible disadvantages for female patients in real-world scenarios.

From the literature review conducted by @Merone_Tsey_Russell_Nagle_2022, women continue to be significantly underrepresented in medical literature, with inadequate reporting and analysis of sex and gender in research and misogynistic perceptions persistently shaping the narrative surrounding these issues. Medical services based on a lack of representative research make diagnoses for diseases and treatments inconvenient, inaccurate, and challenging. The experiences with the medical system could be one of the reasons contributing to female participants reporting higher occurrences of feeling both mentally and physically unwell. The result in this paper again highlights the disparity, reminding us that We still have a long way to acknowledge and solve gender and minority disparity problems in the research field.

## Weaknesses and Future Directions

There are many limitations and weaknesses in the analysis. For instance, counted days of feeling physically and mentally unwell in the past 30 days might not fully capture an individual's mental and physical health status; although feeling unwell in a few days, the severity of the unwellness was not measured. An individual could feel seriously unwell physically or mentally, which could also contribute to a factor for analyzing. On the other hand, no mental health issues other than depression were studied in the paper; different kinds of mental health issues could contribute to physical unwellness in different ways, further research could include those variables for a better narrative of the relationship.

As mentioned before, the variables that are used for the main analysis were only available in the years 2002, 2006, 2010, 2014, 2018, and 2022; the gap between years in the dataset fails to illustrate the trends as its narrative lacks data from the other years. Moreover, not enough education on mental health could also influence the variables we collected, based on the fact that the variables are collected in a self-report manner. Participants might report inaccurate information due to not having mental health-related knowledge. On the other hand, the disparity in gender and other minority groups could influence the variable validity as those groups are more likely to be underrepresented in medical research. There is still much room for improvement.

Lastly, by posterior predictive check, the posterior distribution indicates that the model is not doing a good job of fitting the data. The check result could be due to limitations of the data mentioned above; furthermore, adding more relevant variables to the linear regression equation could as well improve the strength of the model.

There are still many mental and physical health factors that we did not touch on to study their relationship in deeper depth. With existing studies and limitations concluded, future studies could be conducted to improve and capture the relationship between mental and physical health more efficiently in the future.

\appendix

# Appendix

## Survey Questions {#sec-ques-details}

\begin{enumerate}
   \item What is your age group?
   \begin{itemize}
      \item Type in a Number
      \item Not Applicable 
   \end{itemize}
   \item Please specify your gender:
   \begin{enumerate}
      \item Male
      \item Female
      \item Prefer not to say
   \end{enumerate}
   \item Would you say your own health, in general, is excellent, good, fair, or poor?
   \begin{enumerate}
      \item Excellent
      \item Good
      \item Fair
      \item Poor
      \item Not Applicable
   \end{enumerate}
   \item Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?
   \begin{itemize}
      \item Type in a Number 
      \item Not Applicable
   \end{itemize}
   \item Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good?:
   \begin{itemize}
      \item Type in a Number 
      \item Not Applicable
   \end{itemize}
   \item Now I would like to ask you some questions about general health conditions. Has a doctor, nurse, or other health professional EVER told you that you had Depression?
   \begin{enumerate}
      \item Yes
      \item No
      \item Not Applicable
   \end{enumerate}
\end{enumerate}

## Model details {#sec-model-details}

### Posterior predictive check

In @fig-ppcheckandposteriorvsprior, we implement a posterior predictive check and compare the posterior with the prior. By definition, the posterior should be able to be used to simulate data similar to the actual data; however, the posterior simulated data barely recollects the characteristics of the dataset from GSS. This means that the model lacks strength in terms of the relationship. The weakness could be a result of data limitations mentioned in [Discussion].

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "The Model Fits Barely to the Data, Hinting A Weak Model for Explaining the Circumstances"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
\break
### Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows lines bouncing around horizontally and has a nice overlap between the chains. @fig-stanareyouokay-2 is a Rhat plot. From the plot, $\hat{R} = 1.05$, which is very close to 1, and ideally no more than 1.1.  These results suggest no problems in the model we built for our paper.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "The convergence of the MCMC algorithm is checked for the model."
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References